### 2020-09-08 paper/paper.tex
introduce literature and data

> "I am aware that the position is temporary. Especially because it is a political position, decided by the administration. If the current administration is out of power, we are automatically dismissed." - Interview with school principal A, August 2019.

@@

In Brazil, responsibility for primary education is delegated to local governments (paschoal_historia_2009). The municipal educational system has increased in relevance over the past few decades, as municipalities took on a larger proportion of students from state-level governments. Figure \ref{fig:enrollment} plots the total number of students in primary education per government level, including private schools. As of 2016, over 25 million students were enrolled in over 115 thousand municipal schools.

\blue{plots/enrollment_dep.pdf}

The local executive branch has sole prerogative over the hiring and firing of teachers and school principals. School principals are considered positions of trust (\textit{cargos de confiança}) and usually appointed by the mayor or secretary of education (brollo_political_2017). Teachers usually enter the public service through a public exam and are eligible for tenure after two years (gatti_formacao_2010) However municipalities increasingly resort to temporary contracts to hire new teachers as budgetary pressures amount. Municipal teachers and school principals are overseen by a local department of education (militao_educacao_2014).

Funding for municipal education relies primarily on federal transfers, the Fundeb, derived from 10 percent of tax revenues for each level of government to education. At the municipal level, 25 percent of the local budget must be allocated to education, and compensatory federal transfers are institutionalized by law to those municipalities which do not reach the target.^[For more details on the Fundeb, see (here)[https://www.fnde.gov.br/index.php/financiamento/fundeb/sobre-o-plano-ou-programa/sobre-o-fundeb].] While municipalities may be audited to verify whether funds are being properly used, personnel decisions and daily operations are fully under municipal discretion (ferraz_corrupting_2012).

\subsection{Local governance and mayoral coalitions}

Brazil is comprised by over 5 thousand municipalities, each with their own city hall and council. Mayors and city councillors (\textit{vereadores}) are democratically elected every four years, with the possibility of reelection for both.^[The last municipal elections were in 2016. These take place ] The executive is responsible for the implementation of social policies such as education, as well as drafting the municipal budget. The city council, on the other hand, oversees legislation and approves the budget for the fiscal year. As noted by souza_reforma_2004, decisions over how to manage education are primarily in the hands of the mayor, but the city councilors exert significant pressure.

In order to win elections and garner support for their campaign, mayors form electoral alliances with local party labels (dantas_eleicoes_2013). These mayoral coalitions, once in government, are an important basis for approving budgets and, more generally, executive control over the legislative chamber. Interviews with city councilors indicate that the legislature is divided into a pro-government (\textit{governo}) and opposition (\textit{oposição}) groups (\textit{bancada}). While these electoral coalitions do not necessarily remain intact once governments are formed, fragmentation is generally on the margins, with mayors swapping defecting legislators for new ones.^[Interview with C, former chief of staff of municipality J. September 2018.]

Interviews with secretaries of education and mayors suggest that city councilors play an important role in staffing decisions. While mayors hold jurisdiction over personnel decisions, there are extensive consultations between mayors and city councilors to decide who becomes the principal of a school, or which teacher remains in a school or leaves. These executive-legislative consultations ultimately determine the allocation (\textit{alotação}) of educational staff, serving as the primary tool through which mayors reward or coopt legislators into supporting them in the city council. In the words of a set of school administrators in the municipality of I:

> School principal: Here, we are invited to work at the school by the department of education, with the [political] candidate, the city councilor...deciding which are the positions they are searching for and appointing people they think have the necessary qualifications.

> Educational counselor: I was also invited to work here, by the city councilor and the department of education.^[Interview with the administrative board of school A., municipality J. August 2019]

These accounts, along with previous case studies of municipalities in Brazil, suggest that city councilors play an important role in nominating staff. To verify these claims in a broader set of municipalities, as well as outlining the research design, I employ extensive administrative and electoral data collected by the Brazilian federal government. In the next section I describe the data, where it is publicly available, as well as the preparation necessary for the set of estimations presented in the study.

\section{Theory}

In this section, I apply a canonical vote-buying model in legislatures \citep{groseclose_1996_buying, banks_2000_buying} to model mayor's strategy to use patronage to buy legislative votes. For any given electoral term, a mayor wants to enact her preferred policy, but to do so needs to pass it through a simple majority in the city council. To ensure victory, the mayor offers appointment positions to voters, in this case, city councilors, who in exchange vote in favor of the bill.

In this setting, passing a policy proposal requires outbidding the opposition, which generates interesting results regarding the optimal size of the coalition and conditions under which we should observe minimal, super and universal coalitions. In particular, the amount of resources that the opposition controls can force the mayor's hands and increase the amount of patronage necessary to pass the bill. We pursue these empirical implications of the model further in our results section.

The amount of public sector positions the mayor offers to these city councilors also responds to the composition of the city council. The intuition is straightforward: if the city councilors are already likely to support the bill, there is no need to offer them additional side payments. Note that there is no claim about a dichotomy of clientelism, in which there is or is not patronage \citep{stokes_brokers_2013, grindle_jobs_2012}, but rather a variation in degrees: how many appointments must the mayor give out in order to govern on her preferred policy.

For the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.

>>>>

\begin{quote}
"I am aware that the position is temporary. Especially because it is a political position, decided by the administration. If the current administration is out of power, we are automatically dismissed." - Interview with school principal A, August 2019.
\end{quote}

@@

In Brazil, responsibility for primary education is delegated to local governments \citet{paschoal_historia_2009}. The municipal educational system has increased in relevance over the past few decades, as municipalities took on a larger proportion of students from state-level governments. Figure \ref{fig:enrollment} plots the total number of students in primary education per government level, including private schools. As of 2016, over 25 million students were enrolled in over 115 thousand municipal schools.

\begin{figure}[h]
    \label{fig:enrollment}
    \includegraphics{plots/descriptive_enrollment.pdf}
    \centering
\end{figure}
% \blue{plots/enrollment_dep.pdf}

% The local executive branch has sole prerogative over the hiring and firing of teachers and school principals. School principals are considered positions of trust (\emph{cargos de confiança}) and usually appointed by the mayor or secretary of education \citet{brollo_political_2017}. Teachers usually enter the public service through a public exam and are eligible for tenure after two years (gatti_formacao_2010) However municipalities increasingly resort to temporary contracts to hire new teachers as budgetary pressures amount. Municipal teachers and school principals are overseen by a local department of education \citet{militao_educacao_2014}.

% Funding for municipal education relies primarily on federal transfers, the Fundeb, derived from 10 percent of tax revenues for each level of government to education. At the municipal level, 25 percent of the local budget must be allocated to education, and compensatory federal transfers are institutionalized by law to those municipalities which do not reach the target.^[For more details on the Fundeb, see \href{https://www.fnde.gov.br/index.php/financiamento/fundeb/sobre-o-plano-ou-programa/sobre-o-fundeb}{here}. While municipalities may be audited to verify whether funds are being properly used, personnel decisions and daily operations are fully under municipal discretion (ferraz_corrupting_2012).

% \subsection{Local governance and mayoral coalitions}

% Brazil is comprised by over 5 thousand municipalities, each with their own city hall and council. Mayors and city councillors (\textit{vereadores}) are democratically elected every four years, with the possibility of reelection for both.^[The last municipal elections were in 2016. These take place ] The executive is responsible for the implementation of social policies such as education, as well as drafting the municipal budget. The city council, on the other hand, oversees legislation and approves the budget for the fiscal year. As noted by souza_reforma_2004, decisions over how to manage education are primarily in the hands of the mayor, but the city councilors exert significant pressure.

% In order to win elections and garner support for their campaign, mayors form electoral alliances with local party labels (dantas_eleicoes_2013). These mayoral coalitions, once in government, are an important basis for approving budgets and, more generally, executive control over the legislative chamber. Interviews with city councilors indicate that the legislature is divided into a pro-government (\textit{governo}) and opposition (\textit{oposição}) groups (\textit{bancada}). While these electoral coalitions do not necessarily remain intact once governments are formed, fragmentation is generally on the margins, with mayors swapping defecting legislators for new ones.^[Interview with C, former chief of staff of municipality J. September 2018.]

% Interviews with secretaries of education and mayors suggest that city councilors play an important role in staffing decisions. While mayors hold jurisdiction over personnel decisions, there are extensive consultations between mayors and city councilors to decide who becomes the principal of a school, or which teacher remains in a school or leaves. These executive-legislative consultations ultimately determine the allocation (\textit{alotação}) of educational staff, serving as the primary tool through which mayors reward or coopt legislators into supporting them in the city council. In the words of a set of school administrators in the municipality of I:

% > School principal: Here, we are invited to work at the school by the department of education, with the [political] candidate, the city councilor...deciding which are the positions they are searching for and appointing people they think have the necessary qualifications.

% > Educational counselor: I was also invited to work here, by the city councilor and the department of education.^[Interview with the administrative board of school A., municipality J. August 2019]

% These accounts, along with previous case studies of municipalities in Brazil, suggest that city councilors play an important role in nominating staff. To verify these claims in a broader set of municipalities, as well as outlining the research design, I employ extensive administrative and electoral data collected by the Brazilian federal government. In the next section I describe the data, where it is publicly available, as well as the preparation necessary for the set of estimations presented in the study.

% \section{Theory}

% In this section, I apply a canonical vote-buying model in legislatures \citep{groseclose_1996_buying, banks_2000_buying} to model mayor's strategy to use patronage to buy legislative votes. For any given electoral term, a mayor wants to enact her preferred policy, but to do so needs to pass it through a simple majority in the city council. To ensure victory, the mayor offers appointment positions to voters, in this case, city councilors, who in exchange vote in favor of the bill.

% In this setting, passing a policy proposal requires outbidding the opposition, which generates interesting results regarding the optimal size of the coalition and conditions under which we should observe minimal, super and universal coalitions. In particular, the amount of resources that the opposition controls can force the mayor's hands and increase the amount of patronage necessary to pass the bill. We pursue these empirical implications of the model further in our results section.

% The amount of public sector positions the mayor offers to these city councilors also responds to the composition of the city council. The intuition is straightforward: if the city councilors are already likely to support the bill, there is no need to offer them additional side payments. Note that there is no claim about a dichotomy of clientelism, in which there is or is not patronage \citep{stokes_brokers_2013, grindle_jobs_2012}, but rather a variation in degrees: how many appointments must the mayor give out in order to govern on her preferred policy.

% For the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.
